### 2013-12-13 inst/doc/Advanced_writing_of_EML.Rmd

@@

#### Overview of the EML heirarchy

The EML metadata language

@@

```{r}
additionalMetadata <- hf205@additionalMetadata # extracted from previous eml file
```

```{r}
dataset <- new("dataset", 
                title = title,
                creator = creator,
                contact = contact,
                pubDate = pubDate,
                associatedParty = other_researchers,
                keywordSet = keywords,
                coverage = coverage,
                methods = methods,
                dataTable = c(dataTable))

@@

@@

This `eml` object we have just created now has the complete R representation of the EML data.  This is the same class of object returned by `eml_read`, and consequently we can use all of the read methods provided to explore or manipulate it.

@@

While it is difficult to create invalid EML using the constructor functions, it is not impossible so it never hurts to check.  As illustrated in the basic tutorial, we can now publish this file automatically to a central repository such as the KNB.

>>>>

```{r include=FALSE, cache=FALSE}
source("~/.knitr_defaults.R")
```

### Overview of the EML heirarchy

A basic knowledge of the components of an EML metadata file is essential to being able to take full advantage of the language. While more complete information can be found in the official schema documentation, here we provide a general overview of commonly used metadata elements most relevant to describing data tables.

This schematic shows each of the metadata elements we will generate.  Most these elements have sub-components (e.g. a 'publisher' may have a name, address, and so forth) which are not shown for simplicity.  Other optional fields we will not be generating in this example are also not shown.

```yaml
- eml
  - dataset
    - creator
    - contact
    - publisher
    - title
    - pubDate
    - keywords
    - abstract 
    - intellectualRights
    - methods
    - coverage
    - dataTable
      - physical
      - attributeList
  - additionalMetadata
```

The EML R package allows us to assemble these pieces bit by bit, allowing us to work either from the top down or bottom up.

Our `data.set` object captures all of the information required to generate the `dataTable` element: the `data.frame`, which will be written to an external CSV file descibed by the `physical` node, and the `attributeList`, containing the column and unit metadata descriptions.

@@

```{r}
library(reml)
```

@@

```{r}
hf205 <- eml_read(system.file("examples", "hf205.xml", package="reml"))
additionalMetadata <- hf205@additionalMetadata # extracted from previous eml file
```

We now have all of the elements we wished to define for the `dataset` metadata. We can assemble them into a `datset` node as follows:

```{r}
dataset <- new("dataset", 
                title = title,
                creator = creator,
                contact = contact,
                pubDate = pubDate,
                associatedParty = other_researchers,
                keywordSet = keys,
                coverage = coverage,
                methods = methods,
                dataTable = c(dataTable))

@@

Note that because a `dataset` can have multiple dataTable elements, we use the same concatenate notation illustrated earlier, `c(dataTable)`, even though only one element is provided.  This lets R known that the resulting object is a `ListOfdataTable`, not an individual `dataTable` file.  This convention should be followed whenever assigning a value to a slot that can take more than one element of that name.

With most of the metadata now in place, we can construct the top-level element:

@@

This call automatically adds a few additional bits of information, such as a unique identifier `packageId` for the metadata file.  This `eml` object we have just created now has the complete R representation of the EML data.  This is the same class of object returned by `eml_read`, and consequently we can use all of the read methods provided to explore or manipulate it.

@@

While it is difficult to create invalid EML using the constructor functions, it is not impossible so it never hurts to check.

As illustrated in the basic tutorial, we can now publish this file automatically to a central repository such as the KNB.
